---
title: "Distribution Analysis - carsDB"
author: "Daniel Chacón Malaver"
date: "`r format(Sys.Date())`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(ggplot2)
options(digits = 3)

# Load dataset
# Make sure the file exists at: data/carsDB.csv
carsDB <- read.csv("data/carsDB.csv")
```

# 1. Overview

This notebook explores relationships between **weight (wt)** and three variables in the *carsDB* dataset:
- Displacement (**disp**)
- Fuel efficiency (**mpg**)
- Gross horsepower (**hp**)

We compute correlations and visualize trends with **ggplot2**.

# 2. Quick data check

```{r}
str(carsDB)
summary(carsDB[c("wt", "disp", "mpg", "hp")])
```

# 3. Correlations

```{r}
c_wt_disp <- cor(carsDB$wt, carsDB$disp, use = "complete.obs")
c_wt_mpg  <- cor(carsDB$wt, carsDB$mpg,  use = "complete.obs")
c_wt_hp   <- cor(carsDB$wt, carsDB$hp,   use = "complete.obs")

data.frame(
  metric = c("wt vs disp", "wt vs mpg", "wt vs hp"),
  correlation = c(c_wt_disp, c_wt_mpg, c_wt_hp)
)
```

# 4. Scatterplots with smoothing

## 4.1 Weight vs Displacement

```{r}
p1 <- ggplot(carsDB, aes(x = disp, y = wt)) +
  geom_point(alpha = 0.8) +
  geom_smooth(se = TRUE) +
  labs(
    title = "Weight vs Displacement",
    x = "Displacement (cu.in.)",
    y = "Weight (1000 lbs)"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
p1
```

**Notes:** Expect a **positive correlation** (heavier cars tend to have larger displacement).

## 4.2 Weight vs MPG

```{r}
p2 <- ggplot(carsDB, aes(x = mpg, y = wt)) +
  geom_point(alpha = 0.8) +
  geom_smooth(se = TRUE) +
  labs(
    title = "Weight vs MPG",
    x = "Miles/(US) gallon",
    y = "Weight (1000 lbs)"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
p2
```

**Notes:** Expect a **negative correlation** (heavier cars are usually less fuel efficient).

## 4.3 Weight vs Horsepower

```{r}
p3 <- ggplot(carsDB, aes(x = hp, y = wt)) +
  geom_point(alpha = 0.8) +
  geom_smooth(se = TRUE) +
  labs(
    title = "Weight vs Gross Horsepower",
    x = "Gross horsepower",
    y = "Weight (1000 lbs)"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
p3
```

**Notes:** Typically **positive correlation**; inspect for possible anomalies/outliers.

# 5. (Optional) Save figures to `results/`

> Create the folder `results/` in your repo before running this chunk.

```{r, eval=FALSE}
if (!dir.exists("results")) dir.create("results", recursive = TRUE)
ggsave("results/wt_disp_plot.png", p1, width = 7, height = 5, dpi = 300)
ggsave("results/wt_mpg_plot.png",  p2, width = 7, height = 5, dpi = 300)
ggsave("results/wt_hp_plot.png",   p3, width = 7, height = 5, dpi = 300)
```

# 6. Session info

```{r}
sessionInfo()
```
